---
title: "SDH inhibition snMultiome RNA & ATAC-seq"
author: "Dakota Nuttall"
date: "2025-06-17"
R-version: 4.3.3
---
###Cardiomyocyte Analysis###

```{r include=FALSE}
set.seed(1234)
library(Seurat)
library(Signac)
library(chromVAR)
library(dplyr)
library(reticulate)
library(hdf5r)
library(scales)
library(ggplot2)
library(openxlsx)
library(readxl)
library(EnsDb.Mmusculus.v79)
library(BSgenome.Mmusculus.UCSC.mm10)
library(org.Mm.eg.db)
library(JASPAR2020)
library(TFBSTools)
library(harmony)
library(pheatmap)
library(EnhancedVolcano)
library(ggrepel)
library(tidyr)
library(gridExtra)
library(ggpubr)
library(VennDiagram)
library(extrafont)

loadfonts(device = "win") # Use device="pdf" or "postscript" if needed

```
####  Load data  ####

```{r echo=TRUE}
#Load integrated file
aggregated_samples <- readRDS("aggregated_global.rds")

#Initial Subset Cardiomyocyte (CM) population
aggregated_Cardiomyocytes <- subset(aggregated_samples, idents = "Cardiomyocytes")


```

###Re-normalize and re-run PCA analysis (repeat after contamination removal)

```{r echo=TRUE}
aggregated_Cardiomyocytes <- SCTransform(aggregated_Cardiomyocytes, vars.to.regress = c("percent.mt","percent.ribo"),vst.flavor = "v2", method = "glmGamPoi", variable.features.n = 5000)

DefaultAssay(aggregated_Cardiomyocytes) <- "SCT"

#Establish new RNA, ATAC and WNN umaps
aggregated_Cardiomyocytes <- RunPCA(aggregated_Cardiomyocytes, npcs = 50)
aggregated_Cardiomyocytes <- RunHarmony(aggregated_Cardiomyocytes, 
                              group.by.vars = "orig.ident",
                              reduction = "pca", assay.use = "SCT", reduction.save = "harmony_pca",project.dim= F)
aggregated_Cardiomyocytes <- FindNeighbors(object = aggregated_Cardiomyocytes, dims = 1:50, reduction = "harmony_pca")
ElbowPlot(aggregated_Cardiomyocytes, reduction = "pca", ndims = 50)
aggregated_Cardiomyocytes <- RunUMAP(aggregated_Cardiomyocytes, dims = 1:50, reduction = "harmony_pca",reduction.name = "umap_RNA", reduction.key = "UMAP")

DefaultAssay(aggregated_Cardiomyocytes) <- "peakunion"

aggregated_Cardiomyocytes <- FindTopFeatures(aggregated_Cardiomyocytes, min.cutoff = 10)
aggregated_Cardiomyocytes<- RunTFIDF(aggregated_Cardiomyocytes,)
aggregated_Cardiomyocytes <- RunSVD(aggregated_Cardiomyocytes)

#Note: Harmony over-corrects for cell type level ATAC data so we stick with the default "lsi" reduction

aggregated_Cardiomyocytes <- RunUMAP(object = aggregated_Cardiomyocytes, reduction = 'lsi', dims = 2:50, reduction.name = "umap_peakunion", reduction.key = "peakunionUMAP_")

DefaultAssay(aggregated_Cardiomyocytes) <- "SCT"
ElbowPlot(aggregated_Cardiomyocytes, ndims = 50, reduction = "harmony_pca")# ~20-30
FindDimNumber(aggregated_Cardiomyocytes,reduction = "lsi") #~45
aggregated_Cardiomyocytes <- FindMultiModalNeighbors(
  object = aggregated_Cardiomyocytes,
  reduction.list = list("harmony_pca", "lsi"), 
    dims.list = list(1:30, 2:45),
  modality.weight.name = c("SCT.weight","peakunion.weight"),
  verbose = TRUE
)
aggregated_Cardiomyocytes <- RunUMAP(
  object = aggregated_Cardiomyocytes,
  nn.name = "weighted.nn",
  assay = "SCT",
  reduction.name = "WNN_umap",
  reduction.key = "WNN_UMAP_",
  verbose = TRUE,
)

clust_colors <- c("darkorange2","mediumorchid2","steelblue3","olivedrab","indianred1","hotpink","darkred","green3","cyan")
treatment_colors <-c("lightskyblue","pink", "steelblue4","red3")  

```
### Check for potential contamination (skip if already preformed)
```{r echo=TRUE}

aggregated_Cardiomyocytes<- FindClusters(aggregated_Cardiomyocytes, graph.name = "wsnn", algorithm = 3, resolution = seq(.1, 1, 0.1))

#Check for potential contamination by using low resolution first 0.1, which only shows the clusters with the most variable gene expression
Idents(aggregated_Cardiomyocytes) <- aggregated_Cardiomyocytes$wsnn_res.0.1
quick_CT_plot(aggregated_Cardiomyocytes, 2,"WNN_umap",legend = F,label = T, font.size = 8)
quick_cond_plot(aggregated_Cardiomyocytes, 2,"WNN_umap",legend = F)
annotation_markers <- c("Myh6","Ttn","Tnnt2","Ryr2", #Cardiomyocytes
                        "Col1a1","Postn","Pdgfra","Gsn","Htra3",#Cardiac Fibroblasts
                        "Cdh5","Fabp4","Pecam1","Kdr",#Endothelial Cells
                        "Rgs5","Kcnj8","P2ry14","Pdgfrb","Vtn",#Mural cells
                        "Mrc1","Cd68","C1qa","Ms4a7","Ccr2","Cx3cr1",#Myeloid Cells
                        "Cd79a","Ikzf3","Ms4a1","Skap1",#Lymphoid Cells
                        "Upk3b","Upk1b","Clu","Dmkn",# Epicardial cells
                        "Kcna1","Kcna2","Cdh19","Plp1")#Schwann cells

DotPlot_SCT(aggregated_Cardiomyocytes, gene_vector = annotation_markers)

contaminated <- Cells(subset(aggregated_Cardiomyocytes, idents = "1") )
low_quality <- readLines("low_quality_BCs.txt")
low_quality <- c(low_quality,contaminated)
writeLines(low_quality, "low_quality_BCs+CM.txt")

aggregated_Cardiomyocytes <- subset(aggregated_Cardiomyocytes, idents = "0") 
## If contamination removed, repeat all steps after the initial subset


```
### Double check for contamination with higher resolution
```{r echo=TRUE}

aggregated_Cardiomyocytes<- FindClusters(aggregated_Cardiomyocytes, graph.name = "wsnn", algorithm = 3, resolution = seq(.1,.7,.1))

Idents(aggregated_Cardiomyocytes) <- aggregated_Cardiomyocytes$wsnn_res.0.7
quick_CT_plot(aggregated_Cardiomyocytes, 2,"WNN_umap",legend = F,label = T, font.size = 8)
quick_cond_plot(aggregated_Cardiomyocytes, 2,"WNN_umap",legend = F)
DotPlot_SCT(seurat_obj = aggregated_Cardiomyocytes, gene_vector = annotation_markers)

### No clear

```
### Annotate and plot clusters (post contamination removal)
```{r echo=TRUE}
Idents(aggregated_Cardiomyocytes) <- aggregated_Cardiomyocytes$wsnn_res.0.7
#Identify transcriptionally unique clusters 
aggregated_Cardiomyocytes <- PrepSCTFindMarkers(aggregated_Cardiomyocytes, assay = "SCT",verbose = T)
DefaultAssay(aggregated_Cardiomyocytes) <- "SCT"
CM_all_DEGs <- FindAllMarkers(aggregated_Cardiomyocytes, only.pos = T, min.pct = 0.3, logfc.threshold = .5, test.use = "MAST", recorrect_umi = F, assay = "SCT" )
table(CM_all_DEGs$cluster)## Note: clusters 1 and 3 appear to relatively fewer distinguishing genes than 2 and 4. So they will be merged with cluster 0 to reduce cluster complexity and further emphasize the differences in 2 and 4.


Cardiomyocytes_cell_types <- c("CM1","CM1","CM2","CM1","CM3") 
aggregated_Cardiomyocytes@active.ident<-factor(aggregated_Cardiomyocytes@active.ident, labels = Cardiomyocytes_cell_types)

Cardiomyocytes_cell_levels <- c("CM1","CM2","CM3")
aggregated_Cardiomyocytes@active.ident<-factor(aggregated_Cardiomyocytes@active.ident, levels =  Cardiomyocytes_cell_levels)

#Plot cells
                                       
quick_CT_plot(aggregated_Cardiomyocytes, 2,"WNN_umap",legend = F,label = T, font.size = 14)
ggsave("Cardiomyocytes_umap_sub.tiff", height = 5,width = 5)
quick_cond_plot(aggregated_Cardiomyocytes, 2,"WNN_umap",legend = F)
ggsave("Cardiomyocytes_umap_cond.tiff", height = 5,width = 5)


```
## Find DEGs once again for Volcano plots and Gene Ontology (see GO script)

```{r echo=TRUE}
#Perform differential gene expression analysis comparing one cluster to all clusters simultaneously to find unique DEGs.
aggregated_Cardiomyocytes <- PrepSCTFindMarkers(aggregated_Cardiomyocytes, assay = "SCT",verbose = T)
DefaultAssay(aggregated_Cardiomyocytes) <- "SCT"
CM_all_DEGs <- FindAllMarkers(aggregated_Cardiomyocytes, only.pos = F, min.pct = 0.05, logfc.threshold = .25, test.use = "MAST", recorrect_umi = F, assay = "SCT" ,return.thresh = 1)
CM_all_DEGs$gene <- sub("\\..*", "", rownames(CM_all_DEGs))
CM_all_DEGs <- CM_all_DEGs %>% group_by(cluster) %>% mutate(BH_p_val_adj = p.adjust(p_val, method = "BH")) 
CM_all_DEGs <- CM_all_DEGs[!grepl("mt-",CM_all_DEGs$gene),]#Mitochondrial RNA is typically removed during nuclear exraction but some can be bound to ER during extraction so it is removed 
CM_all_DEGs <- as.data.frame(CM_all_DEGs)
CM_all_DEGs_sig <- subset(CM_all_DEGs, BH_p_val_adj < 0.05)
write.xlsx(CM_all_DEGs_sig,"CM_all_DEGs_sig.xlsx")

##For CM1
DefaultAssay(aggregated_Cardiomyocytes) <- "SCT"
CM1_marker <- subset(CM_all_DEGs, cluster == "CM1")
row.names(CM1_marker) <- CM1_marker$gene
CM1_marker_sig <- subset(CM_all_DEGs_sig, cluster == "CM1" )
row.names(CM1_marker_sig) <- CM1_marker_sig$gene

lab_italics <- paste0("italic('", rownames(CM1_marker), "')")
select_labels <-c(head(row.names(subset(CM1_marker, avg_log2FC > 0)),n=15),head(row.names(subset(CM1_marker, avg_log2FC < 0)),n=15))
select_labels <- paste0("italic('",c(head(row.names(subset(CM1_marker, avg_log2FC > 0)),n=15),head(row.names(subset(CM1_marker, avg_log2FC < 0)),n=15)), "')")

#find number of significant upregulated and downregulated genes

up_down_volcano(CM1_marker, selectLab =  select_labels,italicLab = lab_italics,up_label = paste("CM1 Upregulated Genes (",sum(CM1_marker_sig$avg_log2FC > 0)," significant)", sep = ""),down_label = paste("CM1 Downregulated Genes (",sum(CM1_marker_sig$avg_log2FC < 0)," significant)", sep = ""),volcano_name = "CM1_volcano.tiff", up_x = max(CM1_marker$avg_log2FC)+1.5,down_x = min(CM1_marker$avg_log2FC)-1 )

##For CM2
DefaultAssay(aggregated_Cardiomyocytes) <- "SCT"
CM2_marker <- subset(CM_all_DEGs, cluster == "CM2")
CM2_marker <- CM2_marker[!duplicated(CM2_marker$gene), ] # "Pakap" gene has duplicate EnsembleIDs
row.names(CM2_marker) <- CM2_marker$gene
CM2_marker_sig <- subset(CM_all_DEGs_sig, cluster == "CM2" )
CM2_marker_sig <- CM2_marker_sig[!duplicated(CM2_marker_sig$gene), ]
row.names(CM2_marker_sig) <- CM2_marker_sig$gene

lab_italics <- paste0("italic('", rownames(CM2_marker), "')")
select_labels <-c(head(row.names(subset(CM2_marker, avg_log2FC > 0)),n=10), "Nppa","Nppb","Sorbs2","Myh7","Ccn2","Tgfb2",head(row.names(subset(CM2_marker, avg_log2FC < 0)),n=10))
select_labels <- paste0("italic('",c(head(row.names(subset(CM2_marker, avg_log2FC > 0)),n=10), "Nppa","Nppb","Sorbs2","Myh7","Ccn2","Tgfb2",head(row.names(subset(CM2_marker, avg_log2FC < 0)),n=10)), "')")

#find number of significant upregulated and downregulated genes

#Note: Check and adjust label position with up_x and down_x as needed
up_down_volcano(CM2_marker, selectLab =  select_labels,italicLab = lab_italics,up_label = paste("CM2 Upregulated Genes (",sum(CM2_marker_sig$avg_log2FC > 0)," significant)", sep = ""),down_label = paste("CM2 Downregulated Genes (",sum(CM2_marker_sig$avg_log2FC < 0)," significant)", sep = ""),volcano_name = "CM2_volcano.tiff", up_x = max(CM2_marker$avg_log2FC)+1.5,down_x = min(CM2_marker$avg_log2FC)-1 )

##For CM3
DefaultAssay(aggregated_Cardiomyocytes) <- "SCT"
CM3_marker <- subset(CM_all_DEGs, cluster == "CM3")
row.names(CM3_marker) <- CM3_marker$gene
CM3_marker_sig <- subset(CM_all_DEGs_sig, cluster == "CM3" )
row.names(CM3_marker_sig) <- CM3_marker_sig$gene

lab_italics <- paste0("italic('", rownames(CM3_marker), "')")
select_labels <-c(head(row.names(subset(CM3_marker, avg_log2FC > 0)),n=10), "Ezh1","Ccn1","Lrrc10","Tcap","Pdk4","Cited2","Ccnd3","Tecrl",head(row.names(subset(CM3_marker, avg_log2FC < 0)),n=10))
select_labels <- paste0("italic('",c(head(row.names(subset(CM3_marker, avg_log2FC > 0)),n=10), "Ezh1","Ccn1","Lrrc10","Tcap","Pdk4","Cited2","Ccnd3","Tecrl",head(row.names(subset(CM3_marker, avg_log2FC < 0)),n=10)), "')")

#find number of significant upregulated and downregulated genes

up_down_volcano(CM3_marker, selectLab =  select_labels,italicLab = lab_italics,up_label = paste("CM3 Upregulated Genes (",sum(CM3_marker_sig$avg_log2FC > 0)," significant)", sep = ""),down_label = paste("CM3 Downregulated Genes (",sum(CM3_marker_sig$avg_log2FC < 0)," significant)", sep = ""),volcano_name = "CM3_volcano.tiff", up_x = max(CM3_marker$avg_log2FC)+1.5,down_x = min(CM3_marker$avg_log2FC)-1 )

#Compare DEG between individual clusters 

CM1_DEG <- Comparison_DEG(aggregated_Cardiomyocytes,target_cluster = "CM1",comparison_clusters = c("CM2","CM3"))

CM2_DEG <- Comparison_DEG(aggregated_Cardiomyocytes,target_cluster = "CM2",comparison_clusters = c("CM1","CM3"))

CM3_DEG <- Comparison_DEG(aggregated_Cardiomyocytes,target_cluster = "CM3",comparison_clusters = c("CM1","CM2"))

#save DEG results
combined_DEGs <- c(CM1_DEG,CM2_DEG,CM3_DEG)
combined_DEGs_df <-  do.call("rbind", combined_DEGs)

write.xlsx(combined_DEGs_df,"CM_CT_compare_DEGs.xlsx",)

#Plot DEGs of interest

#CM2
genes_of_interest <-c("Nppa","Nppb","Sorbs2","Myh7","Ccn2","Tgfb2")
plot_DEGs(srt = aggregated_Cardiomyocytes,gene_list = genes_of_interest,comp_list = CM2_DEG,target_clust = "CM2",compare_clust = c("CM1","CM3"),x.font = 32)

#CM3
genes_of_interest <-c("Lifr","Ezh1","Ccn1","Lrrc10","Tcap","Pdk4","Cited2","Ccnd3","Tecrl","Foxo3","Sesn1","Foxo1")
plot_DEGs(srt = aggregated_Cardiomyocytes,gene_list = genes_of_interest,comp_list = CM3_DEG,target_clust = "CM3",compare_clust = c("CM1","CM2"),x.font = 34)


```
## Treatment comparisons with MAST
```{r echo=TRUE}
#Perform differential expression analysis between treatment Sham and MI treatments using MAST

CM_SalShamvMalSham  <- FindMarkers(aggregated_Cardiomyocytes, ident.1 = "Saline_Sham", ident.2 = "Malonate_Sham",group.by = "orig.ident",only.pos = F, min.pct = 0.05, logfc.threshold = .25, test.use = "MAST", recorrect_umi = F, assay = "SCT",)
    CM_SalShamvMalSham$BH_p_val_adj <- p.adjust(CM_SalShamvMalSham$p_val, method='BH')
    CM_SalShamvMalSham$gene <- rownames(CM_SalShamvMalSham)
    CM_SalShamvMalSham <- CM_SalShamvMalSham[!grepl("mt-",CM_SalShamvMalSham$gene),]
    CM_SalShamvMalSham$comparison <- "Saline Sham vs Malonate Sham"
CM_SalShamvMalSham_sig <- subset(CM_SalShamvMalSham, BH_p_val_adj < 0.05 )

lab_italics <- paste0("italic('", rownames(CM_SalShamvMalSham), "')")
select_labels <-c(head(row.names(subset(CM_SalShamvMalSham, avg_log2FC > 0)),n=15),head(row.names(subset(CM_SalShamvMalSham, avg_log2FC < 0)),n=10))
select_labels <- paste0("italic('",c(head(row.names(subset(CM_SalShamvMalSham, avg_log2FC > 0)),n=15),head(row.names(subset(CM_SalShamvMalSham, avg_log2FC < 0)),n=10)), "')")

up_down_volcano(CM_SalShamvMalSham, selectLab =  select_labels,italicLab = lab_italics,up_label = paste("Saline Sham Upregulated Genes (",sum(CM_SalShamvMalSham_sig$avg_log2FC > 0)," significant)", sep = ""),down_label = paste("Malonate Sham upregulated Genes (",sum(CM_SalShamvMalSham_sig$avg_log2FC < 0)," significant)", sep = ""),volcano_name = "CM_SalShamvMalSham_volcano.tif", up_x = max(CM_SalShamvMalSham$avg_log2FC)+1.5,down_x = min(CM_SalShamvMalSham$avg_log2FC)-1.5,cols = c("pink2","lightskyblue3","grey" ))

CM_SalMIvMalMI <- FindMarkers(aggregated_Cardiomyocytes, ident.1 = "Saline_MI", ident.2 = "Malonate_MI",group.by = "orig.ident",only.pos = F, min.pct = 0.05, logfc.threshold = .25, test.use = "MAST", recorrect_umi = F, assay = "SCT",)
    CM_SalMIvMalMI$BH_p_val_adj <- p.adjust(CM_SalMIvMalMI$p_val, method='BH')
    CM_SalMIvMalMI$gene <- rownames(CM_SalMIvMalMI)
    CM_SalMIvMalMI <- CM_SalMIvMalMI[!grepl("mt-",CM_SalMIvMalMI$gene),]
    CM_SalMIvMalMI$comparison <- "Saline MI vs Malonate MI"
CM_SalMIvMalMI_sig <- subset(CM_SalMIvMalMI, BH_p_val_adj < 0.05 )

lab_italics <- paste0("italic('", rownames(CM_SalMIvMalMI), "')")
select_labels <-c(head(row.names(subset(CM_SalMIvMalMI, avg_log2FC > 0)),n=8),head(row.names(subset(CM_SalMIvMalMI, avg_log2FC < 0)),n=10))
select_labels <- paste0("italic('",c(head(row.names(subset(CM_SalMIvMalMI, avg_log2FC > 0)),n=8) ,head(row.names(subset(CM_SalMIvMalMI, avg_log2FC < 0)),n=10)), "')")

up_down_volcano(CM_SalMIvMalMI, selectLab =  select_labels,italicLab = lab_italics,up_label = paste("Saline MI Upregulated Genes (",sum(CM_SalMIvMalMI_sig$avg_log2FC > 0)," significant)", sep = ""),down_label = paste("Malonate MI upregulated Genes (",sum(CM_SalMIvMalMI_sig$avg_log2FC < 0)," significant)", sep = ""),volcano_name = "CM_SalMIvMalMI_volcano.tif", up_x = max(CM_SalMIvMalMI$avg_log2FC)+1,down_x = min(CM_SalMIvMalMI$avg_log2FC)-1,cols = c("red3","steelblue4","grey" ))

#Save combined data
combined_CM_DESeq <- rbind(CM_SalShamvMalSham_sig,CM_SalMIvMalMI_sig)
write.xlsx(combined_CM_DESeq,"CM_Sal_vs_Mal_sig.xlsx")

```
## Identify upregulated ATAC peaks

```{r echo=TRUE}
annotation <- GetGRangesFromEnsDb(ensdb = EnsDb.Mmusculus.v79)
annotation <- renameSeqlevels(annotation, mapSeqlevels(seqlevels(annotation), "UCSC"))
DefaultAssay(aggregated_Cardiomyocytes) <- 'peakunion'
Annotation(aggregated_Cardiomyocytes) <- annotation

library(ChIPseeker)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)

annotated_peaks <- annotatePeak(
  peak = granges(aggregated_Cardiomyocytes[["peakunion"]]),
  TxDb = TxDb.Mmusculus.UCSC.mm10.knownGene,
  tssRegion = c(-2000, 2000),
  annoDb = "org.Mm.eg.db"
)

annotated_peaks_df <- as.data.frame(annotated_peaks)

# Add peak column to markers dataframe
annotated_peaks_df$peak <- gsub(
  ":|-",
  "-",
  paste0(
    annotated_peaks_df$seqnames,
    "-",
    annotated_peaks_df$start,
    "-",
    annotated_peaks_df$end
  )
)


# We will then perform differential analysis using linear regression to analyze ATAC Peaks as it can take into account the variance in nCount_ATAC and TSS.Enrichment between conditions

CM1_peaks <- Comparison_peaks(aggregated_Cardiomyocytes,target_cluster = "CM1",comparison_clusters = c("CM2","CM3"),annotated_peaks_df = annotated_peaks_df)
CM2_peaks <- Comparison_peaks(aggregated_Cardiomyocytes,target_cluster = "CM2",comparison_clusters = c("CM1","CM3"),annotated_peaks_df = annotated_peaks_df)
CM3_peaks <- Comparison_peaks(aggregated_Cardiomyocytes,target_cluster = "CM3",comparison_clusters = c("CM1","CM2"), annotated_peaks_df = annotated_peaks_df)

#save results
combined_peaks <- c(CM1_peaks,CM2_peaks,CM3_peaks)
combined_peaks_df <-  do.call("rbind", combined_peaks)

write.xlsx(combined_peaks_df,"CM_CT_compare_peaks.xlsx",)

#identify peaks that are nearby genes of interest
CM3vCM1 <-data.frame(CM3_peaks$CM1$peak,CM3_peaks$CM1$gene_name,CM3_peaks$CM1$BH_p_val_adj, CM3_peaks$CM1$distance)
CM3vCM1

#Repeat for each comparison

plot_peaks(srt = aggregated_Cardiomyocytes,nearest_genes = c("Pdk4","Ccnd3"),peaks_of_interest = c("chr6-5482155-5482792","chr17-47504903-47505217"),comp_list = CM3_peaks,target_clust = "CM3",compare_clust = c("CM1","CM2"), abbrev = "CM",x.font = 34)
```
## Motif analysis
```{r echo=TRUE}
#load motifs 
pfm <- getMatrixSet(
  x = JASPAR2020,
  opts = list(collection = "CORE", tax_group = 'vertebrates', all_versions = FALSE)
)

# first compute the GC content for each peak
aggregated_Cardiomyocytes <- RegionStats(aggregated_Cardiomyocytes, genome = BSgenome.Mmusculus.UCSC.mm10)

# add motifs to CM dataset

aggregated_Cardiomyocytes <- AddMotifs(
  object = aggregated_Cardiomyocytes,
  genome = BSgenome.Mmusculus.UCSC.mm10,
  pfm = pfm,
)

##Analyze motif activity
aggregated_Cardiomyocytes <- RunChromVAR(
  object = aggregated_Cardiomyocytes,
  genome = BSgenome.Mmusculus.UCSC.mm10,
  assay = "peakunion"
)

DefaultAssay(aggregated_Cardiomyocytes) <- "chromvar"

da_peakunion_CM <- FindAllMarkers(
  object = aggregated_Cardiomyocytes,
  only.pos = TRUE,
  mean.fxn = rowMeans,
  fc.name = "avg_diff", logfc.threshold = .25
)

da_peakunion_CM$motif.names <-  ConvertMotifID(aggregated_Cardiomyocytes,assay = "peakunion",id =row.names(da_peakunion_CM))

da_peakunion_CM<- na.omit(da_peakunion_CM)
da_peakunion_CM$BH_p_val_adj <- p.adjust(da_peakunion_CM$p_val, method='BH')
da_peakunion_CM <- subset(da_peakunion_CM, BH_p_val_adj < 0.05)
write.xlsx(da_peakunion_CM, "Cardiomyocytes_motifs.xlsx")

###Plot top CM1 motifs
Motif_IDs <- head(da_peakunion_CM$gene[da_peakunion_CM$cluster == "CM1"],n=6)
Motif_names <- head(da_peakunion_CM$motif.names[da_peakunion_CM$cluster == "CM1"],n=6)

MotifPlot(
  object = aggregated_Cardiomyocytes,
  motifs = row.names(da_peakunion_CM[Motif_IDs,]), assay = "peakunion"
)
ggsave("CM1_top_motifs.tif", height = 5,width = 8)

##Visualize individual motifs with violin plots and umaps
CM1_motifs<-Comparison_motif(aggregated_Cardiomyocytes,target_cluster = "CM1",comparison_clusters = c("CM2","CM3"))

index<- c(1:length(Motif_names))
for(i in index){

Comparison_violin_motif(seurat_obj = aggregated_Cardiomyocytes,target_cluster = "CM1",comparison_clusters = c("CM2","CM3"),feat = Motif_IDs[i],comparison_list = CM1_motifs,title = Motif_names[i],x.font.size = 36)

ggsave(filename = paste("CM_",Motif_names[i],"_Motif_vln.tiff", sep = ""), height = 5, width = 5)

p <- FeaturePlot(
  object = aggregated_Cardiomyocytes,
  features = Motif_IDs[i],
  pt.size = 2,
  reduction = "WNN_umap",
  min.cutoff = 0, cols = c("lightgrey", "darkred")
) + labs(title = Motif_names[i])+
  theme(text=element_text(size=24),legend.text  = element_text( size = 16), axis.text=element_blank(),axis.ticks=element_blank(), axis.line = element_line(linewidth =  1))+ xlab(NULL) + ylab(NULL) +theme(axis.text=element_blank(),axis.title=element_text(size=20),legend.text  = element_text( size = 16),plot.title = element_text(face = "plain",size = 40))
p
ggsave(filename = paste("CM_",Motif_names[i],"_Motif_umap.tiff", sep = ""),plot = p, height = 5, width = 5)
}

#Plot TF RNA expression
DefaultAssay(aggregated_Cardiomyocytes) <- "SCT"
genes_of_interest <- sapply(Motif_names, capitalize_first, USE.NAMES = FALSE)
plot_DEGs(srt = aggregated_Cardiomyocytes,gene_list = genes_of_interest,comp_list = CM1_DEG,target_clust = "CM1",compare_clust = c("CM2","CM3"),x.font = 34)

###Plot top CM2 motifs
DefaultAssay(aggregated_Cardiomyocytes) <- "chromvar"
Motif_IDs <- head(da_peakunion_CM$gene[da_peakunion_CM$cluster == "CM2"],n=6)
Motif_names <- head(da_peakunion_CM$motif.names[da_peakunion_CM$cluster == "CM2"],n=6)
Motif_names <- gsub(pattern = "::", replacement = "-", x = Motif_names)

MotifPlot(
  object = aggregated_Cardiomyocytes,
  motifs = row.names(da_peakunion_CM[Motif_IDs,]), assay = "peakunion"
)
ggsave("CM2_top_motifs.tif", height = 5,width = 8)

#Plot TF RNA expression
DefaultAssay(aggregated_Cardiomyocytes) <- "SCT"
genes_of_interest <- sapply(c("BATF3","BATF","FOS","JUNB","SMAD2","SMAD3","JUN"), capitalize_first, USE.NAMES = FALSE)
plot_DEGs(srt = aggregated_Cardiomyocytes,gene_list = genes_of_interest,comp_list = CM2_DEG,target_clust = "CM2",compare_clust = c("CM1","CM3"),x.font = 34)

##Visualize individual motifs with violin plots and umaps
DefaultAssay(aggregated_Cardiomyocytes) <- "chromvar"
CM2_motifs<-Comparison_motif(aggregated_Cardiomyocytes,target_cluster = "CM2",comparison_clusters = c("CM1","CM3"))

index<- c(1:length(Motif_names))
for(i in index){

Comparison_violin_motif(seurat_obj = aggregated_Cardiomyocytes,target_cluster = "CM2",comparison_clusters = c("CM1","CM3"),feat = Motif_IDs[i],comparison_list = CM2_motifs,title = Motif_names[i],x.font.size = 36)

ggsave(filename = paste("CM_",Motif_names[i],"_Motif_vln.tiff", sep = ""), height = 5, width = 5)

p <- FeaturePlot(
  object = aggregated_Cardiomyocytes,
  features = Motif_IDs[i],
  pt.size = 2,
  reduction = "WNN_umap",
  min.cutoff = 0, cols = c("lightgrey", "darkred")
) + labs(title = Motif_names[i])+
  theme(text=element_text(size=24),legend.text  = element_text( size = 16), axis.text=element_blank(),axis.ticks=element_blank(), axis.line = element_line(linewidth =  1))+ xlab(NULL) + ylab(NULL) +theme(axis.text=element_blank(),axis.title=element_text(size=20),legend.text  = element_text( size = 16),plot.title = element_text(face = "plain",size = 40))
p
ggsave(filename = paste("CM_",Motif_names[i],"_Motif_umap.tiff", sep = ""),plot = p, height = 5, width = 5)
}



###Plot CM3 motifs of interest (main figure)
DefaultAssay(aggregated_Cardiomyocytes) <- "chromvar"
Motif_IDs <- c('MA0052.4','MA0660.1','MA0497.1','MA0773.1','MA0852.2','MA1103.2','MA0480.1','MA0157.2')
Motif_names <- c("MEF2A","MEF2B","MEF2C","MEF2D","FOXK1","FOXK2","FOXO1","FOXO3")

MotifPlot(
  object = aggregated_Cardiomyocytes,
  motifs = row.names(da_peakunion_CM[Motif_IDs,]), assay = "peakunion"
)
ggsave("CM3_top_motifs.tif", height = 5,width = 8)

##Visualize individual motifs with violin plots and umaps
CM3_motifs<-Comparison_motif(aggregated_Cardiomyocytes,target_cluster = "CM3",comparison_clusters = c("CM1","CM2"))


index<- c(1:length(Motif_names))
for(i in index){

Comparison_violin_motif(seurat_obj = aggregated_Cardiomyocytes,target_cluster = "CM3",comparison_clusters = c("CM1","CM2"),feat = Motif_IDs[i],comparison_list = CM3_motifs,title = Motif_names[i],x.font.size = 36)

ggsave(filename = paste("CM_",Motif_names[i],"_Motif_vln.tiff", sep = ""), height = 5, width = 5)

p <- FeaturePlot(
  object = aggregated_Cardiomyocytes,
  features = Motif_IDs[i],
  pt.size = 2,
  reduction = "WNN_umap",
  min.cutoff = 0, cols = c("lightgrey", "darkred")
) + labs(title = Motif_names[i])+
  theme(text=element_text(size=24),legend.text  = element_text( size = 16), axis.text=element_blank(),axis.ticks=element_blank(), axis.line = element_line(linewidth =  1))+ xlab(NULL) + ylab(NULL) +theme(axis.text=element_blank(),axis.title=element_text(size=20),legend.text  = element_text( size = 16),plot.title = element_text(face = "plain",size = 40))
p
ggsave(filename = paste("CM_",Motif_names[i],"_Motif_umap.tiff", sep = ""),plot = p, height = 5, width = 5)
}

#Plot TF RNA expression
DefaultAssay(aggregated_Cardiomyocytes) <- "SCT"
genes_of_interest <- sapply(Motif_names, capitalize_first, USE.NAMES = FALSE)
plot_DEGs(srt = aggregated_Cardiomyocytes,gene_list = genes_of_interest,comp_list = CM3_DEG,target_clust = "CM3",compare_clust = c("CM1","CM2"),x.font = 34)

#Save motif data
combined_motifs <- c(CM1_motifs,CM2_motifs,CM3_motifs)
combined_motifs_df <-  do.call("rbind", combined_motifs)
combined_motifs_df$motif.names <-  ConvertMotifID(aggregated_Cardiomyocytes,assay = "peakunion",id =combined_motifs_df$motif.ID)

write.xlsx(combined_motifs_df,"CM_CT_compare_motifs.xlsx",)

motif_list <- list("CM all" = da_peakunion_CM, "CM comparisons" = combined_motifs_df) 

# Write the entire list to an Excel file
write.xlsx(motif_list, "Supplemental table 7 (motifs).xlsx")
  
```
## Proportion plots and subcluster heatmaps
```{r echo=TRUE}

aggregated_Cardiomyocytes$Cell_state <- Idents(aggregated_Cardiomyocytes)
#These functions can be found in the helper functions script
proportion_plots(aggregated_Cardiomyocytes, cond_color = treatment_colors, clust_color = clust_colors, name.clust = "CM_prop_sub.tif", name.cond = "CM_prop_cond.tif",height = 4, width = 4)

prop_heatmap(aggregated_Cardiomyocytes, "CM_heat_prop.tif")

#Save dataset
saveRDS(aggregated_Cardiomyocytes,"Cardiomyocytes.rds")
```

